---
title: "Exploring Statistics with R"
author: "Matthias Bannert"
date: "11 Dec 2015"
output: ioslides_presentation

---

## Likert Scale Visualization Example

```{r,echo=FALSE,message=FALSE,fig.height=2}
library(likert)
library(reshape2)

load("../data/pisa_tabbed.RData")
ptab <- acast(tab, ST04Q01 ~ ST62Q13, value.var = "Percentage")
ddat <- data.frame(Item = rownames(ptab), ptab)

# plot it with the likert package
likert.bar.plot(likert(summary = ddat), center = 2) + ggtitle("Declarative Fraction") + 
  theme_bw()

```

## Histograms with ggplot2

```{r,echo=FALSE,message=FALSE}
require(ggplot2)
# read our big5 data set
big5 <- read.csv2("../data/BIG5/data.csv",sep="\t")

# let's select some countries... 
# mention the :: syntax
big5$country <- plyr::revalue(big5$country,
                              c("US" = "USA",
                                "DE" = "Germany",
                                "CA" = "Canada",
                                "CN" = "China",
                                "ES" = "Spain",
                                "FR" = "France")
)

# set up the ggplot object
countries <- ggplot(data = big5) 

# the point geom
# countries + geom_point(aes(country,age))

# maybe we should remove ugly outliers...
big5 <- subset(big5 ,age < 120)
countries <- ggplot(data = big5) 

# the layer by layer thing... 
c_h <- countries +
  geom_histogram(aes(x = age,
                     y = ..density..),
                 binwidth=1,
                 colour="black",
                 fill="white")
# add a density
library(ggthemes)
c_h_d  <- c_h +
  geom_density(aes(x = age),
               alpha = .2,fill = "green") + theme_bw()

c_h_d_m <- c_h_d + 
  geom_vline(aes(xintercept = mean(age, na.rm=T)),
             color="red", linetype="dashed", size=1) +
  theme_bw()

c_h_d_m

library(ggmap)
world_map <- map_data("world")

library(dplyr)

grpd_cntry <- group_by(na.omit(big5),country)
aggregated <- summarize(grpd_cntry,
          E1_mean = mean(E1))

full_countries <- merge(aggregated,world_map,by.x = "country",
                        by.y = "region",all.y = T)

full_countries_sel <- merge(aggregated,world_map,by.x = "country",
                        by.y = "region")


```

## Maps with ggplot2
```{r,echo=FALSE}
info <- ggplot(full_countries, aes(map_id = country))
info + 
  geom_map(aes(fill = E1_mean),
           map = world_map,
           color ="white") + 
  expand_limits(x = world_map$long, y = world_map$lat) +
  scale_fill_gradient(low="black", high="blue") +
  theme_bw() + theme_tufte() +
  coord_map(xlim=c(-180,180))

```



## Maps with GoogleVis

```{r setOptions, message=FALSE,echo=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

```{r ComboExample, results='asis', tidy=FALSE,echo=FALSE}
## Add the mean
unq_info <- unique(full_countries_sel[,c("country","E1_mean")])
unq_info$country <- droplevels(unq_info$country)
levels(unq_info$country)[6] <- "United States"

G <- gvisGeoChart(unq_info, "country", "E1_mean", 
                  options=list(width=400, height=400))
plot(G)
```

## Overview

- Repetition
- regular expressions 
- HWA
- Some advanced plots (ggplot2)
- Factor analysis
- solving assignments

# Basic Data Types
```{r,eval=FALSE}
# vector
a_vector <- c(1,3,45,43)
# matrix
m <- matrix(1:4,nrow = 2)
# matrix
dframe <- data.frame(a = c(1,2,3), b=c(3,2,1))
# list
list(a_vector,m,dframe)
```

## Basic Commands

```{r,eval=FALSE}
# install libs, load libs
install.packages("ggplot2")
library(ggplot2)
# load,save,read,write
load("somefile.RData")
save("object1","object2",file="somefile.RData")
# workspace
ls()
rm(list=ls())
# data
head()
tail()
summary()
table()
c()
t()
mean()
sd()
var()
# ... 
```

## Apply / Lapply
```{r,eval= FALSE}
# apply function to every element of a list
lapply(l,class)
# row wise anonymous function
apply(x,1,function(x) sum(x)-1)
```


## Basic Plots
```{r,eval=FALSE}
# scatterplot
plot(x,y)
# histogram
hist(x)
# boxplot
boxplot(x ~ somecategory)
# pairwise scatterplot
pairs(dataframe)
```

## Simulate Data 
```{r,eval=FALSE}
rnorm(1000)
rbinom(100,2,.5)
rexp(...)
rgamma(...)
rcauchy(...)
rt(...)
# ...
```

## Regressions and Testing
```{r,eval=FALSE}
# basic fit formula syntax
fit1 <- lm(y~x1 + x2,data)
# standard output
summary(fit1)

# tests
t.test(x,y)
t.test(x ~ categorical_var)
library(lmtest)
# heterosked.
bptest()

```

## knitr / markdown

```{r,eval=FALSE}
htmlreg(fit1)
texreg(fit1)
kable(dataframe)

```

<!-- break !!-->

## HAVE a BREAK 

## Regular Expression

- matching based on patterns
- available in many programming languages
- pattern definition can become complex but is very powerful
- great resources: [Regxr](http://regexr.com/), [regex101](https://regex101.com)


## HWA

- questions, problems? 


## ggplot2 

- grammar of graphics / layer by layer
- advanced plotting library for pdf, png etc. 
- most popular among non-interactive charts
- interactive charts: googleVis, rCharts, ggvis, ... 
(plotting most often done with javascript)
- [Hadley Wickham](http://had.co.nz/) (R Studio chief data scientist)
- great documentation: [ggplot2](http://docs.ggplot2.org/current/)
- ggplot2 book: [ggplot2 book](http://ggplot2.org/book/)



















