---
title: "Exploring Statistics with R"
author: "Matthias Bannert"
date: "8 Nov 2015"
output: ioslides_presentation
---

## Task A: Load the Data

Load EITHER the mtcars dataset (easy), the mlb dataset (medium) or
the KOF Businuess situation (hard) into R.

a) mtcars: data(mtcars)
b) mlb: Ilias mlb.RData
c) KOF:
https://datenservice.kof.ethz.ch/api/sets/ds_bs_monthly?mime=csv


## Task A C'd: Basic Report

* Which variables are in the dataset?
* What's the size (rows,cols) of your dataset?
* Which is the most/least economic car (1), who is the tallest/shortest player (2), which is the best/worth month in total (3)?
* Provide some summary statistics such as means quantiles etc?
* What about the correlation structure of your data set? Inspect visually. What can you see? (hint: see pairs(),cor() and image())
* Form an interesting sub groups and compare means in these subgroups.
(hint: factor variables are good starting point)


## Getting Efficient
- functions
- anymous functions (functions without a name)
- the apply family

## Parts of a Function
- name
- body
- arguments 

Besides: try to understand the concept of **scope** from the
very beginning. 

<!-- Let's build a basic function -->


## The Apply Family
- lapply
- sapply
- apply (beware of matrix coercion!!)
- ... 

Idea: Run a function on every element of the other argument.

<!-- Lapply in action -->

<!-- Comparing Groups
- t.test
- boxplot with notches
- the formula syntax
-->

## A Quick Glimpse at R's lm Function
- part of stats package (comes with core installation)
- formula syntax
- data parameter works well with data.frames
- returns object of class lm
- comprehensive summary method
- see also glm() for the generalized linear model

## Math Refresher: OLS coefficients
$$
  Y = X\beta + e \\
  e = Y - X\beta \\
  e'e = (Y' - \beta'X')(Y - X\beta) \\
  = Y'Y - Y'X\beta - \beta'X'Y + \beta'X'X\beta \\
  = Y'Y - 2Y'X\beta + \beta'X'X\beta \\
  \frac{\partial e'e}{\partial \beta} = -2X'Y + 2\beta X'X
$$

set to zero:
$$
  2X'Y = 2\beta X'X \\
  \beta = (X'X)^{-1}X'Y
$$

## Implementation: What Operations do we need? 

- transpose
- matrix multiplication
- invert matrix / solve equation

## Start a Custom Linear Model Fit Function 
- make sure input is numeric
- add an intercept to the data
- what is the input (arguments)?
- what will the output look like?
- what is a suitable name for the function (naming conventions)?


<!-- Saturday Lunch -->

## Weedsmoking in Utopia
Let's consider the following admittedly far fetched example in which grade point average is only determined by a handful of variables. Note that some of these variables are not directly observable to the researcher.

- weedsmoking: binary indicator for regular weed consumption
- weedsmokingstar: grams consumed
- parents: binary indicator for alarming parental background
- lazy: laziness score
- sport: sportiness score
- distance: distance to Amsterdam from where student grows up (instrument)


## Utopia Potheads: The True Model
$$
  gpa = 7.8 - 0.2*weedsmoking - 0.2 * parents \\
        - 0.15 * sport - 0.4 * lazy + u
$$

and 

$$
weedsmokingstar = -3.5+2.5*parents-0.25*sport + \\
  1.2*lazy-0.01*distance+error 
$$

with weedsmoking = 0, if weedsmokingstar < 6, else weedsmoking = 1.


## Texreg 
```{r,echo=FALSE,message=FALSE}
library(texreg)
fit1 <- lm(mpg~am+vs,data=mtcars)
fit2 <- lm(mpg~am+vs+wt,data=mtcars)
```

```{r,results='asis',warning=FALSE,message=FALSE,echo=FALSE}
htmlreg(list(fit1,fit2))
```










